PhD Thesis

### 2016-10-03 main.tex
Misc changes to introduction

% \section{}
Efficient queries are written either as declarative database queries that are
optimized by database optimizers, or as manually optimized programs using
collection libraries. Compare maintaining anagraphic data for a population
as either a relational database or a collection:
\begin{itemize}
\item If the population data is maintained as a database, a programmer that needs to
  look up quickly people by age, can just add an index and keep the lookup query
  unchanged; the index is then automatically maintained.
\item If the population data is maintained as an in-memory collection, adding an
  index requires modifying by hand queries to reuse it for maximum performance,
  and ensuring all changes to population data update the index to maintain
  correctness. This is error-prone and violations are common \pg{XXX}, even if
  there is quite some research on the problem.
\end{itemize}
Modern libraries offer high-level APIs for manipulating (in-memory) collections;
equivalently, such an API is an embedded domain-specific language (EDSL).
Collection operations are typically executed directly; however, they can instead
construct a query representation that can then be optimized.

@@

\section{Our design for incrementalizing queries}
We designed our incrementalization system by abstracting from needs for
collection APIs.
\begin{itemize}
\item collection APIs are typically higher-order. This allows further
  flexibility compared to first-order query languages: in particular we can
  design new operators in terms of existing ones.
\item many collection types support equations that can be used for optimization.
  For instance
\end{itemize}

>>>>

A program manipulating collections of data will often perform queries on those
collections. These collections can be big, so query execution can be expensive
and warrant considerable optimization effort.
% \section{}
Efficient queries are written either as declarative database queries that are
optimized by database optimizers, or as manually optimized programs using
collection libraries. Compare maintaining anagraphic data for a population
as either a relational database or a collection:
\begin{itemize}
\item If the population data is maintained as a database, a programmer that needs to
  look up quickly people by age, can just add an index and keep the lookup query
  unchanged; the index can then be automatically maintained.
\item If the population data is maintained as an in-memory collection, adding an
  index requires modifying by hand queries to reuse it for maximum performance,
  and ensuring all changes to population data update the index to maintain
  correctness. This is error-prone and violations are common \pg{XXX}, even if
  there is quite some research on the problem.
\end{itemize}
Modern libraries offer high-level APIs for manipulating (in-memory) collections;
equivalently, such an API is an embedded domain-specific language (EDSL).
Compared to SQL, an EDSL will typically have first-class support for using
arbitrary user-defined functions in queries, because it will allow .

\pg{Examples!}

Collection operations are typically executed directly; however, they can instead
construct a query representation that can then be optimized.

@@

\section{Our design for incrementalizing queries}
We designed our incrementalization system by abstracting from needs and ideas for
collection APIs.
\begin{itemize}
\item collection APIs are typically higher-order. This allows further
  flexibility compared to first-order query languages: in particular we can
  design new operators in terms of existing ones.
\item many collection types support equations that can be used for optimization.
  For instance
\end{itemize}
